 

### 2022-05-23 thesis/abstract.qmd
Improve minor issues in thesis

Die maximale Sauerstoffaufnahme (V̇O~2max~) ist die wichtigste physiologische Kenngröße der Ausdauerleistungsfähigkeit. Zur Bestimmung der V̇O~2max~ werden während eines Rampentests die Atemgase kontinuierlich gemessen. Die üblicherweise im breath-by-breath-Verfahren erhobenen Daten erfordern eine Weiterverarbeitung um die V̇O~2max~ zu bestimmen. Unterschiedliche Verfahren der Datenverarbeitung können die gemessenen V̇O~2max~ beeinflussen. Ziel der vorliegenden Arbeit ist es, anhand eines scoping review zu bestimmen, welche Datenverarbeitungsstrategien in der wissenschaftlichen Literatur genutzt werden, um diese im Anschluss anhand experimenteller Daten zu vergleichen. Für den Review wurde eine Zufallsstichprobe von 500 Artikeln untersucht; der experimentelle Vergleich erfolgte anhand von 72 standardisierten Rampentests auf dem Laufband. Die im Review untersuchten Studien benutzen eine Vielzahl von verschiedenen Datenverabeitungsstrategien, welche häufig nur unzureichend beschrieben werden. Die unterschiedlichen Strategien können die V̇O~2max~ im Median zu mehr als 5% beeinflussen, mit deutlich höheren individuellen Abweichungen. Die Ergebnisse verdeutlichen den Bedarf einer Standardisierung und korrekter Darstellung der benutzten Datenverarbeitungsstrategien zur Bestimmung der V̇O~2max~. Diese Arbeit bietet hierfür Empfehlungen zur Verwendung und Vergleichbarkeit verschiedener Strategien, sowie Richtlinien zur Beschreibung dieser im Rahmen wissenschaftlicher Studien.

>>>>

Die maximale Sauerstoffaufnahme (V̇O~2max~) ist die wichtigste physiologische Kenngröße der Ausdauerleistungsfähigkeit. Zur Bestimmung der V̇O~2max~ werden während eines Rampentests die Atemgase kontinuierlich gemessen. Die üblicherweise im breath-by-breath-Verfahren erhobenen Werte erfordern eine Weiterverarbeitung um die V̇O~2max~ zu bestimmen. Unterschiedliche Verfahren der Datenverarbeitung können die gemessene V̇O~2max~ beeinflussen. Ziel der vorliegenden Arbeit ist es, anhand eines scoping review zu bestimmen, welche Datenverarbeitungsstrategien in der wissenschaftlichen Literatur genutzt werden, um diese im Anschluss anhand experimenteller Daten zu vergleichen. Für den Review wurde eine Zufallsstichprobe von 500 Artikeln untersucht; der experimentelle Vergleich erfolgte anhand von 72 standardisierten Rampentests auf dem Laufband. Die im Review untersuchten Studien benutzen eine Vielzahl von unterschiedlichen Datenverabeitungsstrategien, welche häufig nur unzureichend beschrieben wurden. Die verschiedenen Strategien können die V̇O~2max~ im Median zu mehr als 5% beeinflussen, bei teils deutlich größeren individuellen Abweichungen. Die Ergebnisse verdeutlichen den Bedarf einer Standardisierung und korrekten Darstellung der benutzten Datenverarbeitungsstrategien zur Bestimmung der V̇O~2max~. Diese Arbeit bietet hierfür Empfehlungen zur Verwendung und Vergleichbarkeit verschiedener Strategien, sowie Richtlinien zur Beschreibung dieser im Rahmen wissenschaftlicher Studien.

### 2022-05-23 thesis/conclusion.qmd
Improve minor issues in thesis

The determination of V̇O~2max~ from raw breath-by-breath data requires some form of data post-processing. As shown by the scoping review, researchers use a variety of strategies for this data processing. I demonstrated on experimental data, that different processing strategies systematically bias the V̇O~2max~ assessment, which can ultimately lead to misclassification of individuals or groups, or to incorrect evaluation of interventions. Despite contradictory recommendations, binned time average remain by far the most common data processing strategy. The calculation intervals for time averages range from 5 to 60 seconds in the reviewed literature, which heavily influences the V̇O~2max~ value obtained. Almost all research did only incompletely report on their data processing to determine V̇O~2max~, and as such can be regarded non-reproducible. To ensure valid and reproducible V̇O~2max~ assessments, researchers should at minimum report their processing strategy based on a set of guidelines I developed in this thesis. Despite calls for standardization of data processing, to date no evidence for an optimal strategy exists, but digital filtering (e.g., zero-phase Butterworth filter) appears to generate a more reasonable data smoothing compared to traditional (moving) averages. I encourage researchers to share their raw gas exchange data and use programming software for data analysis, to minimize the bias due to data processing when determining the V̇O~2max~.

>>>>

The determination of V̇O~2max~ from raw breath-by-breath data requires some form of data post-processing. As shown by the scoping review, researchers use a variety of strategies for this data processing. I demonstrated on experimental data, that different processing strategies systematically bias the V̇O~2max~ assessment, which can ultimately lead to misclassification of individuals or groups, or to incorrect evaluation of interventions. Despite contradictory recommendations, binned time averages remain by far the most common data processing strategy. The calculation intervals for time averages range from 5 to 60 seconds in the reviewed literature, which heavily influences the V̇O~2max~ value obtained. Almost all research did only incompletely report on their data processing to determine V̇O~2max~, and as such can be regarded as non-reproducible. To ensure valid and reproducible V̇O~2max~ assessments, researchers should at minimum report their processing strategy based on a set of guidelines I developed in this thesis. Despite calls for standardization of data processing, to date no evidence for an optimal strategy exists, but digital filtering (e.g., zero-phase Butterworth filter) appears to generate a more reasonable data smoothing compared to traditional (moving) averages. I encourage researchers to share their raw gas exchange data and use programming software for data analysis, to minimize the bias due to data processing when determining the V̇O~2max~.

### 2022-05-23 thesis/discussion.qmd
Improve minor issues in thesis

The collected data shows, that current research uses a variety of strategies to determine V̇O~2max~, which directly influences the values obtained. Many articles only incompletely report on their methods and use outdated strategies. These practices hinder the validity and reproducibility of V̇O~2max~ measurement.

@@

Despite calls to use moving averages [@robergs2003; @robergs2010], binned time averages remain the most commonly used data processing strategy to determine V̇O~2max~ in the reviewed literature (see @fig-strategies). These numbers are generally in line with the findings of the non-systematic search by Midgley et al. @midgley2007 and the survey by Robergs et al. @robergs2010. It is somewhat surprising, that practices have not changed in recent years despite the publication of recommendations, that discourage researchers from using binned averages [@robergs2010]. Using binned time averages leads to systematically lower V̇O~2max~ values as compared to moving averages (see @fig-comparison). The peak in oxygen uptake may fall in between two averaging intervals, resulting in an underestimation of V̇O~2max~ (see @fig-example for an example). Binned time average revoke the most important argument for measuring breath-by-breath: the high temporal resolution of data. Despite these arguments speaking against the use of binned time averages, my review demonstrates that they remain extremely common in the scientific literature.

Breath-based averages seem to be more common (≈ 8%) than reported previously (\< 1%) [@midgley2007], but less common than assessed in self-reporting (≈ 17%) [@robergs2010]. The increasing proportion of breath-based averages may be explained by publications in the recent years advocating for their use [@robergs2003; @robergs2010]. However the proportion of breath-based averages is still much smaller than that of binned time averages, which have not been recommended in such a way.

@@

The exact reasons for exercise scientist to ignore most recommendations by using binned time averages remain unknown. Some researchers may simply not be aware of the impact different data processing strategies have on the V̇O~2max~. But past publications on this issue have been widely cited [@robergs2010; @midgley2007; @astorino2009] and should be known to most scientists in this field. Researchers may also use binned averages for traditional reasons. Douglas bags, as well as many mixing chamber devices, measure the oxygen uptake over fixed time intervals, producing data apparently similar to those by a binned time average of breath-by-breath data. But comparability with older data should only be an issue when using data acquired by different measurement methods within one analysis. Current studies using new breath-by-breath data do not reasonably need to rely on outdated methods of data processing.

A major source for choosing suboptimal processing strategies may be limitations by analysis software. My review shows that most researchers use the software of the metabolic cart's manufacturer to analyse the gas exchange data. These software may by default output binned time averages instead of raw breath-by-breath data. Moreover, further processing (e.g., interpolation, moving averages) may require the use of additional software. This may also explain why digital filtering has --- despite recommended by Robergs et al. @robergs2010 --- has not been used in a single study reviewed here: standard distributions of common data analysis software (e.g., Microsoft Excel) lack the capability to perform such operations. The common malpractices of data processing are likely to be both attributed to a lack of awareness and a lack of easy-to-use software solutions.

@@

The different data processing strategies found in the literature directly bias the V̇O~2max~ determination (see @fig-comparison), and as such can influence the classification of individuals, the evaluation of training success, and the assessment of V̇O~2max~ attainment. In accordance with previous findings [@johnson1998; @astorino2009; @sell2021] --- and pure logic --- longer calculation intervals lead to lower V̇O~2max~ values (see @fig-comparison). The analysed data shows median differences up to 5-7% between processing strategies, which is in accordance with previous research [@martin-rincon2019]. Some studies reported even higher mean differences of up to 20% [@myers1990], but only when including raw breath-by-breath data in the comparison. The evaluation of unprocessed raw data for its maximum is highly erroneous (see the individual breath data points in @fig-example) and as such is not performed in research (see @fig-duration); so there is no reason to include it in a comparison to other strategies. While previous research was often conducted in sedentary or recreational individuals, this thesis now presents evidence, that a similar effect of data processing strategies on V̇O~2max~ exists in well-trained athletes.

Binned time averages lead to systematically lower V̇O~2max~ values compared with moving averages, for the reasons explained above. While this general trend has been acknowledged previously [@martin-rincon2019], it has not been quantified. The data presented in this thesis suggest a ≈-1% lower median V̇O~2max~ when using binned averages compared with moving averages of the same calculation interval length. This difference is well within the measurement error of most if not all metabolic carts, but it is systematic and as such may bias the evaluation in scenarios where small changes in V̇O~2max~ are important (e.g. in elite sports).

@@

Digital filtering seems to be a promising method, as it markedly reduces the variability moving averages have due to single data outliers (see @fig-example). I therefore agree with Robergs et al. @robergs2010, who recommend digital filters over any classical averaging procedure. The exact type of filter and values of filter parameters to be used have not yet been systematically investigated, and it is unclear whether it is possible at all to determine criteria for identifying an optimal filter. However, a Butterworth filter, as proposed by Robergs et al. @robergs2010, seems to produce reasonably smoothed data. Note, that a single Butterworth filter comprises a time lag, a fact highlighted by Weir et al. @weir2004 that was not acknowledged by Robergs et al. @robergs2010. To allow for a correct phasing of data, as well as for a correct V̇O~2max~ determination when the measurement is terminated shortly after reaching exhaustion, a zero-phase filter is needed. A zero-phase forward-backward Butterworth filter (as suggested by Weir et al. @weir2004) seems to produce reasonable smoothing when used with the parameters suggested by Robergs et al. @robergs2010 (low-pass cut-off frequency: 0.04 Hz; third order; see @fig-example). It should be noted, that despite using the same parameters the degree of filtering varies from that by Robergs et al. @robergs2010, as the filter is applied twice, which changes the overall magnitude of parameters. Other filtering parameters may be as reasonable as this, but it is unclear how to objectively investigate this issue.

@@

@@

The results of this thesis suggest that comprehensive reporting facilitates approximate comparisons of V̇O~2max~ data on a group level derived using different data processing strategies. But on an individual level and for a precise comparison, reporting is not enough, as differences between data processing strategies vary between individuals and are potentially influenced by training status. Sharing of the raw metabolic gas exchange data is the solution to this problem. It allows any researcher to recalculate the V̇O~2max~ using their preferred data processing strategy. Anonymization of raw gas exchange data files should not be a hinder sharing, since most of these files are structured in a simple way, which allows to easily remove any personal information (if this had not been done in the metabolic cart's system before). In terms of reproducibility of V̇O~2max~ determination, a way even superior to correct reporting and raw data sharing is to additionally share data analysis code. This allows any researcher to independently reproduce the V̇O~2max~ processing conducted within a study, but requires the data analysis to take place in a programming (or at least code generating) environment. Luckily such programs for the purpose of analyzing gas exchange data exist as free open-source software [@maturana2022; @nolte2022].

>>>>

This thesis aimed to map current practices of data processing strategies to determine V̇O~2max~ and to compare them on experimental data. The results show, that current research uses a variety of strategies to determine V̇O~2max~, which directly influences the values obtained. Many articles only incompletely report on their methods and use outdated strategies. These practices hinder the validity and reproducibility of V̇O~2max~ measurement.

@@

Despite calls to use moving averages [@robergs2003; @robergs2010], binned time averages remain the most commonly used data processing strategy to determine V̇O~2max~ in the reviewed literature (see @fig-strategies). These numbers are generally in line with the findings of the non-systematic search by Midgley et al. @midgley2007 and the survey by Robergs et al. @robergs2010. It is somewhat surprising, that practices have not changed in recent years despite the publication of recommendations, that discourage researchers from using binned averages [@robergs2010]. Using binned time averages leads to systematically lower V̇O~2max~ values as compared to moving averages (see @fig-comparison). The peak in oxygen uptake may be attained between two averaging intervals, resulting in an underestimation of V̇O~2max~ (see @fig-example for an example). Binned time average revoke the most important argument for measuring breath-by-breath: the high temporal resolution of data. Despite these arguments speaking against the use of binned time averages, my review demonstrates that they remain extremely common in the scientific literature.

Breath-based averages seem to be more common (\~8%) than reported previously (\<1%) [@midgley2007], but less common than assessed in self-reporting (\~17%) [@robergs2010]. The increasing proportion of breath-based averages may be explained by publications in the recent years advocating for their use [@robergs2003; @robergs2010]. However the proportion of breath-based averages is still much smaller than that of binned time averages, which have not been recommended in such a way.

@@

The exact reasons for exercise scientist to ignore most recommendations by using binned time averages remain unknown. Some researchers may simply not be aware of the impact different data processing strategies have on the V̇O~2max~. But past publications on this issue have been frequently cited [@robergs2010; @midgley2007; @astorino2009] and should be known to most scientists in this field. Researchers may also use binned averages for traditional reasons. Douglas bags, as well as many mixing chamber devices, measure the oxygen uptake over fixed time intervals, producing data apparently similar to those by a binned time average of breath-by-breath data. But comparability with older data should only be an issue when using data acquired by different measurement methods within one analysis. Current studies using new breath-by-breath data do not reasonably need to rely on outdated methods of data processing.

A major source for choosing suboptimal processing strategies may be limitations by analysis software. My review shows that most researchers use the software of the metabolic cart's manufacturer to analyse the gas exchange data. These software may by default output binned time averages instead of raw breath-by-breath data. Moreover, further processing (e.g., interpolation, moving averages) may require the use of additional software. This may also explain why digital filtering has --- despite recommended by Robergs et al. @robergs2010 --- not been used in a single study reviewed here: standard distributions of common data analysis software (e.g., Microsoft Excel) lack the capability to perform such operations. The common malpractices of data processing are likely to be both attributed to a lack of awareness and a lack of adequate software solutions.

@@

The different data processing strategies found in the literature directly bias the V̇O~2max~ determination (see @fig-comparison), and as such influence the classification of individuals, the evaluation of training success, and the assessment of V̇O~2max~ attainment. In accordance with previous findings [@johnson1998; @astorino2009; @sell2021] --- and pure logic --- longer calculation intervals lead to lower V̇O~2max~ values (see @fig-comparison). The analysed data shows median differences up to 5-7% between processing strategies, which is in accordance with previous research [@martin-rincon2019]. Some studies reported even higher mean differences of up to 20% [@myers1990], but only when including raw breath-by-breath data in the comparison. The evaluation of unprocessed raw data for its maximum is highly erroneous (see the individual breath data points in @fig-example) and as such is not performed in research (see @fig-duration); so there is no reason to include it in a comparison to other strategies. While previous research was often conducted in sedentary or recreational individuals, this thesis now presents evidence, that a similar effect of data processing strategies on V̇O~2max~ exists in trained athletes.

Binned time averages lead to systematically lower V̇O~2max~ values compared with moving averages, for the reasons explained above. While this general trend has been acknowledged previously [@martin-rincon2019], it has not been quantified. The data presented in this thesis suggest a \~1% lower median V̇O~2max~ when using binned averages compared with moving averages of the same calculation interval length. This difference is well within the measurement error of most if not all metabolic carts, but it is systematic and as such may bias the evaluation in scenarios where small changes in V̇O~2max~ are important (e.g. in elite sports).

@@

Digital filtering seems to be a promising method, as it markedly reduces the variability moving averages have due to single data outliers (see @fig-example). I therefore agree with Robergs et al. @robergs2010, who recommend digital filters over any classical averaging procedure. The exact type of filter and values of filter parameters to be used have not yet been systematically investigated, and it is unclear whether it is possible to determine criteria for identifying an optimal filter. However, a Butterworth filter, as proposed by Robergs et al. @robergs2010, seems to produce reasonably smoothed data. Note, that a single Butterworth filter comprises a time lag, a fact highlighted by Weir et al. @weir2004 that was not acknowledged by Robergs et al. @robergs2010. To allow for a correct phasing of data, as well as for a correct V̇O~2max~ determination when the measurement is terminated shortly after reaching exhaustion, a zero-phase filter is needed. A zero-phase forward-backward Butterworth filter (as suggested by Weir et al. @weir2004) seems to produce reasonable smoothing when used with the parameters suggested by Robergs et al. @robergs2010 (low-pass cut-off frequency: 0.04 Hz; third order; see @fig-example). It should be noted, that despite using the same parameters the degree of filtering varies from that by Robergs et al. @robergs2010, as the filter is applied twice, which changes the overall magnitude of parameters. Other filtering parameters may be as reasonable as this, but it is unclear how to objectively investigate this issue.

@@

The results of this thesis suggest that comprehensive reporting facilitates approximate comparisons of V̇O~2max~ data on a group level derived using different data processing strategies. But on an individual level and for a precise comparison, reporting is not enough, as differences between data processing strategies vary between individuals and are potentially influenced by training status. Sharing of the raw metabolic gas exchange data is the solution to this problem. It allows any researcher to recalculate the V̇O~2max~ using their preferred data processing strategy. Anonymization of raw gas exchange data files should not be a hinder for sharing, since most of these files are structured in a simple way, which allows to easily remove any personal information (if this had not been done in the metabolic cart's system before). In terms of reproducibility of V̇O~2max~ determination, a way even superior to correct reporting and raw data sharing is to additionally share data analysis code. This allows any researcher to independently reproduce the V̇O~2max~ processing conducted within a study, but requires the data analysis to take place in a programming (or at least code generating) environment. Luckily such programs for the purpose of analyzing gas exchange data exist as free open-source software [@maturana2022; @nolte2022].

@@

### 2022-05-23 thesis/introduction.qmd
Improve minor issues in thesis

V̇O~2max~ is defined as the highest rate a body can consume oxygen [@bassett2000]. Determined by measuring gas exchange data, it is one of the most common assessed exercise parameters in sports science and medicine. V̇O~2max~ highly corresponds with endurance performance in heterogeneous groups [@reaburn2008; @costill1973; @tanaka1990] and can be regarded the most relevant physiological parameter for predicting endurance performance [@bassett2000]. Many training programs aim to target the V̇O~2max~, and studies evaluate the quality of interventions by measuring changes in V̇O~2max.~

Researchers predominantly measure V̇O~2max~ during exercise tests to exhaustion. In a laboratory setting such tests commonly take place on treadmills or cycling ergometer. Fatigue prior to reaching the V̇O~2max~ results in underestimating its true value. Therefore most exercise protocols consist of a continuous or stepwise increase in load with an optimal duration of 8-12 minutes [@buchfuhrer1983 ; @yoon2007]. While these narrow limits have been challenged, considerably shorter or longer protocols should not be used [@midgley2008]. Despite protocol type and protocol duration, factors such as motivation [@midgley2018], exercise modality [@myers1991] and biological variability [@katch1982] bias the V̇O~2max~ determination.

@@

In absence of one commonly accepted method, data processing varies among the literature. In light of the influence on outcome variables, many articles highlighted the need to report processing strategies in research articles [@myers1990; @astorino2009; @scheadler2017; @johnson1998]. Midgley et al. @midgley2007 were the first to evaluate reported data processing strategies for breath-by-breath analyses in selected journals. They found that almost all studies reporting their methods choose binned time averaging, with only 1 in 117 using a moving time and a moving breath average, respectively. One third of the studies did not describe their processing method at all. While providing interesting first insights into reporting and processing practices, the search by Midgley et al. @midgley2007 was not systematic and its methods were not described in detail. Robergs et al. @robergs2010 wrote that to investigate the current state of data processing strategies, two possible approaches are *"(i) a summary of published research, and (ii) a survey circulated via the Internet to as many exercise physiologists as possible"*. They provided the latter one with a total of 75 respondents, who reported a great variety in data processing strategies. Most researchers reported the use of binned time averages over 30 or 60 seconds. Astonishingly, about half of the respondents admitted that their data processing strategy was rather chosen due to subjective influences as opposed to objective reasoning. Practices in reporting and processing may have changed more than a decade after the publication of such numbers.

@@

Selected data processing strategies have been extensively compared in the literature. Due to the absence of a systematic mapping of current practices, these works lacked the reasoning of which strategies to compare. Many studies compared different averaging intervals, but not averaging types (e.g. moving breath vs. binned time) [@astorino2009; @midgley2007]. @martin-rincon2019 Martin-Rincon et al. provided formulas for comparing data processing strategies by investigating a data set of sedentary or recreational trained athletes, using two different metabolic carts. Therefore in this work motivation and measurement devices may have interacted with the influence of processing strategies.

>>>>

V̇O~2max~ is defined as the highest rate a body can consume oxygen [@bassett2000]. Determined by measuring gas exchange data, it is one of the most commonly assessed exercise parameters in sports science and medicine. V̇O~2max~ highly corresponds with endurance performance in heterogeneous groups [@reaburn2008; @costill1973; @tanaka1990] and can be regarded the most relevant physiological parameter for predicting endurance performance [@bassett2000]. Many training programs aim to target the V̇O~2max~, and studies evaluate the quality of interventions by measuring changes in V̇O~2max.~

Researchers predominantly measure V̇O~2max~ during exercise tests to exhaustion. In a laboratory setting such tests commonly take place on treadmills or cycling ergometer. Fatigue prior to reaching the V̇O~2max~ results in underestimating its true value. Therefore most exercise protocols consist of a continuous or stepwise increase in load with an optimal duration of 8-12 minutes [@buchfuhrer1983 ; @yoon2007]. While these narrow limits have been challenged, considerably shorter or longer protocols should not be used [@midgley2008]. Despite protocol type and protocol duration, factors such as motivation [@midgley2018], exercise modality [@myers1991] and biological variability [@katch1982] affect the V̇O~2max~ determination.

@@

In absence of one commonly accepted method, data processing varies among the literature. In light of the influence on outcome variables, many articles highlighted the need to report processing strategies in research articles [@myers1990; @astorino2009; @scheadler2017; @johnson1998]. Midgley et al. @midgley2007 were the first to evaluate reported data processing strategies for breath-by-breath analyses in selected journals. They found that almost all studies reporting their methods choose binned time averaging, with only 1 in 117 using a moving time and a moving breath average, respectively. One third of the studies did not describe their processing method at all. While providing interesting first insights into reporting and processing practices, the search by Midgley et al. @midgley2007 was not systematic and its methods were not described in detail. Robergs et al. @robergs2010 wrote that to investigate the current state of data processing strategies, two possible approaches are *"(i) a summary of published research, and (ii) a survey circulated via the Internet to as many exercise physiologists as possible"*. They provided the latter one with a total of 75 respondents, who reported a great variety in data processing strategies. Most researchers reported the use of binned time averages over 30 or 60 seconds. Astonishingly, about half of the respondents admitted that their data processing strategy was rather chosen due to subjective influences as opposed to objective reasoning. Practices in reporting and processing may have changed more than a decade after the publication of these numbers.

@@

Selected data processing strategies have been extensively compared in the literature. Due to the absence of a systematic mapping of current practices, these works lacked the reasoning of which strategies to compare. Many studies compared different averaging intervals, but not averaging types (e.g. moving breath vs. binned time) [@astorino2009; @midgley2007]. Martin-Rincon et al. @martin-rincon2019 provided formulas for comparing data processing strategies by investigating a data set of sedentary or recreational trained athletes, using two different metabolic carts. Therefore in this work motivation and measurement devices may have interacted with the influence of processing strategies.

### 2022-05-23 thesis/methods.qmd
Improve minor issues in thesis

@@

The search results from both data bases were merged and checked for the presence of a Digital Object Identifier (DOI). Entries without DOI were excluded to allow for automated removal of duplicates by DOI matching in the next step. After removing the duplicates I conducted an automated title scanning to exclude results that were likely no original research articles. All titles that contained one of the following words were excluded: review, correction, meta-analysis, comment, retraction, editorial, erratum, reply.

@@

A total of N = 72 exercise tests were analysed for this study. Due to a miscalculation, the preregistration had incorrectly stated a number of 76 tests. The data were from previous research on the metabolic profile of endurance runners [@quittmann_unpubl_a; @quittmann_unpubl_b]. The tested individuals were experienced distance runners (15 female, 54 male; three of the males participated in both studies). The V̇O~2max~ tests were conducted in March to September 2019 [@quittmann_unpubl_a] and March to October 2021 [@quittmann_unpubl_b] using an identical exercise protocol. Participants run on a treadmill (saturn 300/100, h/p/cosmos sports & medical 127 GmbH, Nussdorf-Traunstein, Germany) with 1% inclination for ten minutes at a velocity of 2.8 m·s^-1^ as a warm-up. After preparing the gas exchange measures, they started a ramp protocol with an initial speed of 2.8 m·s^-1^ for two minutes and subsequently increased velocity by 0.15 m·s^-1^ every 30 seconds. The researchers provided verbal encouragement and terminated the exercise when the participants reached subjective exhaustion.

@@

To compare different processing methods within and between individuals, I choose to express the V̇O~2max~ normalized to a reference procedure. The reference procedure was chosen as being the most commonly applied in current literature as determined by the systematic review. Individual V̇O~2max~ values were expressed in reference to this procedure, where a value of 1 means that the processing method yields exactly the same V̇O~2max~ value as the reference method. I calculated the data for all integer parameter values within the range of the values found in the literature during the review. On a group level, I calculated the median and 10%- and 90%-quantiles of each processing strategy.

>>>>

The search results from both data bases were merged and checked for the presence of a Digital Object Identifier (DOI). Entries without DOI were excluded to allow for automated removal of duplicates by DOI matching in the next step. After removing the duplicates I conducted an automated title scanning to exclude results that were likely no original research articles. All titles that contained one of the following words were excluded: *'review', 'correction', 'meta-analysis', 'comment', 'retraction', 'editorial', 'erratum', 'reply'.*

@@

@@

A total of N = 72 exercise tests were analysed for this study. Due to a miscalculation, the preregistration had incorrectly stated a number of 76 tests. The data were from previous research on the metabolic profile of endurance runners [@quittmann_unpubl_a; @quittmann_unpubl_b]. The tested individuals were experienced distance runners (15 female, 54 male; three of the males participated in both studies). The V̇O~2max~ tests were conducted in March to September 2019 [@quittmann_unpubl_a] and March to October 2021 [@quittmann_unpubl_b], respectively, while using an identical exercise protocol. Participants run on a treadmill (saturn 300/100, h/p/cosmos sports & medical 127 GmbH, Nussdorf-Traunstein, Germany) with 1% inclination for 8 minutes at a velocity of 2.8 m·s^-1^ as a warm-up. After preparing the gas exchange measures, they started a ramp protocol with an initial speed of 2.8 m·s^-1^ for two minutes and subsequently increased velocity by 0.15 m·s^-1^ every 30 seconds. The researchers provided verbal encouragement and terminated the exercise when the participants reached subjective exhaustion.

@@

To compare different processing methods within and between individuals, I chose to express the V̇O~2max~ normalized to a reference procedure. The reference procedure was chosen as being the most commonly applied in current literature as determined by the systematic review. Individual V̇O~2max~ values were expressed in reference to this procedure, where a value of 1 means that the processing method yields exactly the same V̇O~2max~ value as the reference method. I calculated the data for all integer parameter values within the range of the values found in the literature during the review. On a group level, I calculated the median and 10%- and 90%-quantiles of each processing strategy.

### 2022-05-23 thesis/results.qmd
Improve minor issues in thesis

Reporting practices of the methodology of gas exchange measures differed widely across the literature (see @tbl-reporting). Almost half (44.2%) of the articles did not report any information regarding their data processing strategy. One in twenty articles (5.8%) provided a rationale for their used strategy. Only a single article [@maturana2021] reported information regarding all the investigated criteria.

@@

The average V̇O~2max~ as determined by a binned 30-second average was 62.2±6.3 ml·min^-1^·kg^-1^ (mean ± standard deviation). Applying different data processing strategies for V̇O~2max~ determination leads to different outcome values (see @fig-comparison).

@@

Binned time averages systematically generate lower V̇O~2max~ values compared to moving averages. Decreasing the calculation interval to 5 seconds leads to a 3-4% median increase of V̇O~2max~ values. Notably on an individual level these increases may be lower (≈2%) or much higher (\>10%) than the median value.

>>>>

Reporting practices of the methodology of gas exchange measures differed widely across the literature (see @tbl-reporting). Almost half (44.2%) of the articles did not report any information regarding their data processing strategy. About one in twenty articles (5.8%) provided a rationale for their used strategy. Only a single article [@maturana2021] reported information regarding all the investigated criteria.

@@

The average V̇O~2max~ as determined by a binned 30-second average was 62.2 ± 6.3 ml·min^-1^·kg^-1^ (mean ± standard deviation). Applying different data processing strategies for V̇O~2max~ determination leads to different outcome values (see @fig-comparison).

@@

Binned time averages systematically generate lower V̇O~2max~ values compared to moving averages. Decreasing the calculation interval to 5 seconds leads to a 3-4% median increase of V̇O~2max~ values. Notably on an individual level these increases may be lower (\~2%) or much higher (\>10%) than the median value.
